A sample computation

### 2024-10-03 draft_systoles.tex
minor misprints corrections

@@

\begin{Cla}
\begin{itemize}
	\item The second cohomology $H^2(G, M)$ classify the extensions $0 \to M \to \tilde G \to G \to 0$;
	\item In particular, $H^2(G, \mathbb{Z}/2\mathbb{Z})$ correspond to the extensions of degree $2$;
\item Given an extension $0 \to M \to \tilde G \to G \to 0$ and a principle $G$-bundle over $X$ we get its class in $H^2(X, M)$ via the boundary (in other words, Bockstein) homomorphism $H^1(X, G) \to H^2(X, M)$.  In details, the Bockstein goes by taking the $1$-cocycle defining the $G-$bundle to $\tilde G$ and apply the Cech differential;
	\item $H^2( \mathbf{SL}_3(\mathbb{R}), \mathbb{Z}/ 2 \mathbb{Z}) \cong
	H^2( \mathbf{SL}_3(\mathbb{F}_2), \mathbb{Z}/ 2 \mathbb{Z}) \cong \mathbb{Z}/2\mathbb{Z}$;
\item The corresponding unique extensions of degree two are $\widetilde{\mathbf{SL}_3}(\mathbb{R})$ and $\mathbf{SL}_2(\mathbb{F}_7)$.
	We make use of the exceptional isomorphism $\mathbf{PSL}_2(\mathbb{F}_7) \cong \mathbf{SL}_3(\mathbb{F}_2)$, see, for example, \cite{merzon_psl_2_7_sl_3_2}.
\item If $E$ is a principle $\mathbf{SL}_3(\mathbb{R})$-bundle on $X$ the corresponding class in $H^2(X, \mathbb{Z}/ 2 \mathbb{Z})$ is called $w_2(E)$, the second Stiefel-Whitney class;
\item If $E$ is a principle $\mathbf{SL}_3(\mathbb{F}_2)$-bundle on $X$ let us denote the corresponding class in $H^2(X, \mathbb{Z}/ 2 \mathbb{Z})$ by $u_2(E)$.
\end{itemize}
\end{Cla}
This is all very well known.
For the cohomology of $\mathbf{SL}_3(\mathbb{F}_2)$ see, for example, \cite{adem2006lecturescohomologyfinitegroups}.

Now, consider the natural mappings
\[
i: \SL \embeds {\mathbf{SL}_3(\mathbb{R})}
, \] 
\[
\pi: \SL \embeds {\mathbf{SL}_3(\mathbb{F}_2)}
.\]

@@

Define $D_1' = 
\left( \begin{smallmatrix}0&6\\1&0\end{smallmatrix}\right) 
, D_2' =
\left( \begin{smallmatrix}2&3\\3&5\end{smallmatrix}\right)
\in \mathbf{SL}_2(\mathbb{F}_7)$.
Project the matrices to 
$\in \mathbf{PSL}_2(\mathbb{F}_7) \cong \in \mathbf{SL}_3(\mathbb{F}_2) $.
Pick a pair of commuting lifts $D_1, D_2 \in \mathbf{SL}_3(\mathbb{Z})$.
One can take them to be 
\begin{equation}
D_1 = 
\begin{pmatrix}
1 & 1 & 0 \\
0 & 1 & 0 \\
0 & 1 & 1
\end{pmatrix}, 
D_2 =
\begin{pmatrix}
1 & 0 & 0 \\
0 & 1 & 0 \\
0 & 1 & 1
\end{pmatrix}
.\end{equation}

@@

By the splitting principle
using that the roots of the characteristic polynomial 
\ref{rem_roots_estimate} 
satisfy \[
\alpha_1, 1 + p \alpha_1 < 0, 
\alpha_2, 1 + p \alpha_2 < 0, 
\alpha_3, 1 + p \alpha_3 > 0 
\] 
we get the Stiefel-Whitney class to be
\[
w(E) = w(E_1) w(E_2) w (E_3)
= (1 + x + y) (1 + x + y) (1) 
= 1
,\]
and thus $w_2(E) = 0$ indeed does vanish.

Modulo $2$ one easily spots that $B = A^5 \pmod{2}$, thus their lifts to $\mathbf{SL}_2(\mathbf{F}_7)$ do commute and the image of the Bockstain morphism vanishes; thus $u_2(E) = 0$.
% the explanation is that we should consider SL_3(F_2) in the automorphisms of F_8 = F_2[t]/t^3 + t^2 + 1 as a linear space over F_2; then A is multiplication by t and B is a multiplication by 1 + t = t^5. 
\end{proof}

>>>>

\section{Regulators}
We mimic the proofs from a Keith Conrad's write-up on the Dirichlet unit theorem \cite{conraddirichlet}.

@@

\begin{Cla}
\begin{itemize}
	\item The second cohomology $H^2(G, M)$ classify the extensions $0 \to M \to \tilde G \to G \to 0$;
	\item In particular, $H^2(G, \mathbb{Z}/2\mathbb{Z})$ correspond to the extensions of index two;
\item Given an extension $0 \to M \to \tilde G \to G \to 0$ and a principle $G$-bundle over $X$ we get its class in $H^2(X, M)$ via the boundary (in other words, Bockstein) homomorphism $H^1(X, G) \to H^2(X, M)$.  In details, the Bockstein goes by taking the $1$-cocycle defining the $G-$bundle to $\tilde G$ and apply the Cech differential;
	\item $H^2( \mathbf{SL}_3(\mathbb{R}), \mathbb{Z}/ 2 \mathbb{Z}) \cong
	H^2( \mathbf{SL}_3(\mathbb{F}_2), \mathbb{Z}/ 2 \mathbb{Z}) \cong \mathbb{Z}/2\mathbb{Z}$;
\item The corresponding unique extensions of index two are $\widetilde{\mathbf{SL}_3}(\mathbb{R})$ and $\mathbf{SL}_2(\mathbb{F}_7)$.
	We make use of the exceptional isomorphism $\mathbf{PSL}_2(\mathbb{F}_7) \cong \mathbf{SL}_3(\mathbb{F}_2)$, see, for example, \cite{merzon_psl_2_7_sl_3_2}.
\item If $E$ is a principle $\mathbf{SL}_3(\mathbb{R})$-bundle on $X$ the corresponding class in $H^2(X, \mathbb{Z}/ 2 \mathbb{Z})$ is called $w_2(E)$, the second Stiefel-Whitney class;
\item If $E$ is a principle $\mathbf{SL}_3(\mathbb{F}_2)$-bundle on $X$ let us denote the corresponding class in $H^2(X, \mathbb{Z}/ 2 \mathbb{Z})$ by $u_2(E)$.
\end{itemize}
\end{Cla}
This is all very well known.
For the cohomology of $\mathbf{SL}_3(\mathbb{F}_2)$ see, for example, \cite{adem2006lecturescohomologyfinitegroups}.

Now, consider the natural mappings
\[
i: \SL \embeds {\mathbf{SL}_3(\mathbb{R})}
, \] 
\[
\pi: \SL \projects {\mathbf{SL}_3(\mathbb{F}_2)}
.\]

@@

Define $D_1' = 
\left( \begin{smallmatrix}0&6\\1&0\end{smallmatrix}\right) 
, D_2' =
\left( \begin{smallmatrix}2&3\\3&5\end{smallmatrix}\right)
\in \mathbf{SL}_2(\mathbb{F}_7)$.
Project the matrices to 
$\in \mathbf{PSL}_2(\mathbb{F}_7) \simeq \mathbf{SL}_3(\mathbb{F}_2) $.
Pick a pair of commuting lifts $D_1, D_2 \in \mathbf{SL}_3(\mathbb{Z})$.
One can take them to be 
\begin{equation}
D_1 = 
\begin{pmatrix}
1 & 1 & 0 \\
0 & 1 & 0 \\
0 & 1 & 1
\end{pmatrix}, 
D_2 =
\begin{pmatrix}
1 & 0 & 0 \\
0 & 1 & 0 \\
0 & 1 & 1
\end{pmatrix}
.\end{equation}

@@

Diagonalizing $A, B$ over reals by the splitting principle
using that the roots of the characteristic polynomial 
\ref{rem_roots_estimate} 
satisfy 
\[
\alpha_1, 1 + p \alpha_1 < 0, 
\] 
\[
\alpha_2, 1 + p \alpha_2 < 0, 
\] 
\[
\alpha_3, 1 + p \alpha_3 > 0 .
\] 
we get the Stiefel-Whitney class to be
\[
w(E) = w(E_1) w(E_2) w (E_3)
= (1 + x + y) (1 + x + y) (1) 
= 1
,\]
and thus $w_2(E) = 0$ indeed does vanish.

Modulo $2$ one easily spots that $B = A^5 \pmod{2}$, thus their lifts to $\mathbf{SL}_2(\mathbf{F}_7)$ do commute and the image of the Bockstain morphism vanishes; thus $u_2(E) = 0$.
% the explanation is that we should consider SL_3(F_2) in the automorphisms of F_8 = F_2[t]/t^3 + t^2 + 1 as a linear space over F_2; then A is multiplication by t and B is a multiplication by 1 + t = t^5. 
\end{proof}
\begin{Rem}
Similar computations show that the pushforward in second homology of $\mathbb{Z}^2 \to \SL$ defined via $A^3, B$ vanishes, too.
\end{Rem}
